### 2020-02-24 Geopyter.tex
Major rewrite incorporating Reviewer feedback. Still more work to do.

\documentclass[letter, 11pt,titlepage]{article}

% Need to do some more reading: http://plain-text.co/pull-it-together.html
% Convert from LaTeX to Markdown:
% pandoc --standalone --wrap=none --base-header-level=2 -f latex -t gfm+smart+citations --filter=pandoc-citeproc --bibliography=Geopyter.bib --csl=apalike.csl -M Title="GeoPyTeR: Mashable (Geography) Teaching Resources for Python" -B header.md Geopyter.tex -o README.md

\newcommand{\gp}{\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}~\/}
\newcommand{\eg}{e.g.~\/}
\newcommand{\ie}{i.e.~\/}
\newcommand{\pysal}{\textsc{p}y\textsc{sal}~\/}
\newcommand{\comment}[1]{\todo[inline, color=green!40]{#1}}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[hyphens]{url}
\usepackage{fancyvrb} % facny verbatim
\usepackage{textcomp}
\usepackage{upquote}  % Turns quotes in verbatim to straight
\usepackage{varioref} % for vpageref[above]
\usepackage{fullpage}
\usepackage{graphicx} 
\usepackage{tabularx}
\newcolumntype{s}{>{\hsize=.25\hsize \raggedright\arraybackslash}X}
\newcolumntype{b}{>{\hsize=.75\hsize}X} 	
\usepackage{booktabs}
\usepackage{wrapfig}
\usepackage[font=small,labelfont=bf]{caption}
%% Sets page size and margins
%\usepackage[letter,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% For citations
%\usepackage{natbib}
\usepackage[%
% numberedbib,
  natbibapa
]{apacite}

%% Useful packages
\usepackage{amsmath}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=false, allcolors=black]{hyperref}
\def\tightlist{}

\title{Geographical Python Teaching Resources: GeoPyTeR}
\author{Jonathan Reades \and Sergio J. Rey}
%\author{}

\begin{document}
\maketitle

\begin{abstract}
Geopyter, an acronym of Geographical Python Teaching Resources, provides
a hub for the distribution of `best practice' in computational and
spatial analytic instruction; enabling instructors to quickly and
flexibly remix contributed content to suit their needs and delivery
framework; and encouraging contributors from around the world to `give
back' whether in terms of how to teach individual concepts or deliver
whole courses. As such, Geopyter is positioned at the confluence of two
powerful streams of thought in software and education: the Free and Open
Source Software (FOSS) movement in which contributors help to build
better software, usually on an unpaid basis, in return for having access
to better tools and the recognition of their peers); and the rise of
Massive Open Online Courses (MOOCs), which seek to radically expand
access to education by moving course content online and providing access
to students anywhere in the world at little or no cost. This paper sets
out in greater detail the origins and inspiration for Geopyter, the
design of the system and, through examples, the types of innovative
workflows that it enables for teachers. We believe that tools like
Geopyter, which build on open teaching practices and promote the
development of a shared understanding of what it is to \emph{be} a
computational geographer represent an opportunity to expand the impact
of this second wave of innovation in instruction while reducing the
demands placed on those actively teaching in this area.
\end{abstract}

\section{Introduction}\label{introduction}

Although \citet{Donoho2017} traces the origins of data science back to
\citeauthor{Tukey1962}'s \textit{The future of data analysis}
(\citeyear{Tukey1962}), it is only recently that this set of previously disparate practices for working with large data sets has begun to be formalised in a way that might allow it to be taught in a university context instead of acquired on the job through `learning-by-doing'. One of the most profound practical impacts of this development has been on the ways in which developers share and execute code: although systems for combining code, commentary, and results have been around for some time (\eg R-Markdown) these were intended primarily for replicating research outputs and not for the teaching of a new technique or writing of an interactive introduction to a software library. The emergence of data science---and the overall pace of change in the tools and programming libraries that it employs---gave new impetus to the search for lightweight ways of sharing code and documentation, and an interactive browser-based technology called Jupyter was the result.

As a result, university educators in many disciplines are playing catch-up with both the methods and the tools. Geography has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago. At the same time, there is growing evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what a trained geographer used to do with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline). However, there remains unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques, so in this article we present a means of bringing the benefits of data science-approaches to instruction using Jupyter `notebooks' to the teaching of Python programming and geospatial analysis: Geographical Python Teaching Resources (GeoPyTer).

\section{Statement of Need}

Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question that teaching students and researchers how to code and how to \textit{think} computationally presents new challenges \citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}). Consequently, the majority of resources used in the classroom are being developed from scratch at each institution, often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}). Not only has this led to the duplication of effort at multiple sites, but it has also had an enormous impact on the productivity of early career researchers, many of whom are in their first teaching post, as commonly measured by universities.

The overarching purpose of this software is therefore to address two seemingly
contradictory issues: the recognition that no two instructors teach in precisely
the same way, and the fact that few instructors have the luxury of both time and
tenure to develop compelling new course material in splendid isolation. This is 
where the Geographical Python Teaching Resource
(\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although
developed with geographers in mind, it provides a generic means by which
instructors in any discipline can selectively incorporate and re-mix programming
and conceptual content from existing Jupyter notebooks while providing their own
`gloss' for this content. Our approach therefore seeks to develop a rich community of developers and teachers who work \textit{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and whoe support instructors in quickly adapting existing materials to new delivery formats.

\section{Origins and Inspirations}\label{origins-and-inspirations}

Like many open source projects, \gp grew out of a need to scratch an itch: in our roles as teachers of (geo)computational concepts and methods we grappled on an annual basis with the demands of developing and updating instructional material with complex interdependencies. And, while the fundamental computing and analytic concepts may remain fairly stable, the field is highly-dynamic in terms of both the content (\ie what should be in the curriculum) and the code (\ie how applications should be taught). Currently, every time a widely-used library package is updated a protracted sequence of revisions ensues: some teachers, pressed for time, may resist making use of---or even alerting students to---the new features; others may leap in enthusiastically, only to discover that they hadn't fully thought through the challenges of conflicting dependencies and evolving feature sets.

We initially explored with others---faculty with whom we regularly
collaborate---the potential to `divide and conquer': we would each focus on
different topics and then combine these contributions to create a full course.
Though received with some enthusiasm, this plan did not survive contact with
reality: first, when actually faced with the uncertain promise of receiving a
full course in return for our efforts, many of us felt more secure developing
our own material; second, we had underestimated the value that we each attach to
the ability to personalise material to our own style of teaching. Since the
shared materials were being developed individually according to the individual's
own `style', any rationalisation or personalisation would have to be done by the
instructor at a later date and without support from the rest of the project.

Going back to the drawing board, we looked again at our own experiences of
successful programming and instructional efforts. In particular, we took note of
two powerful streams of thought in software and education: the importance
attached by the Free and Open Source Software (\textsc{foss}) movement to both
peer-recognition \citep{raymond_cathedral_1999} and distributed version control;
and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive
Open Online Courses (\textsc{mooc}s) with their use of rich, online learning
experiences (\eg \citealp{Trafford2011,Cabiria2012}). Reflecting on our initial
failure, we decided against trying to supply a single set of polished course
materials and instead looked for a way to provide teachers with building blocks
from which they could assemble an \textit{individualised} course suited to their
institutional and pedagogical needs.

We drew inspiration from the \textsc{foss} iPython project
\citep{perez_2007_ipython} and its transformative impact on the sharing of
integrated code, text, and multimedia resources. And we also looked to the
Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project
\cite{rey_19sea} in which contributors take ownership of application areas
through managed collaboration instead of a contributory `free-for-all'. The
subsequent emergence of the Jupyter project \citep{kluyver16} from iPython has
led to the rapid growth of a rich ecosystem of tools that support the kinds of
interaction needed to make this an effective teaching resource:
multiuser-notebooks (jupyterhub), automation of grading (nbgrader), interactive
widgets for visualisation (ipywidgets), and packages to facilitate manipulation
of the notebook itself (nbformat and notebook).

With these criteria and resources in mind, we began looking for ways to lower
the `entry costs' for faculty to both take from, and share into, the corpus of
instructional material, as well as for a way to address the
customisation/localisation challenge. \gp was our solution: a single, open hub
of teaching concepts and practical programming tasks that can be flexibly
assembled into classes, or entire courses! In the \gp approach we recognise that
teachers need to be able to develop their own classroom `story' at their own
pace (and in conformance with their institution's teaching pattern), while
supporting them with a library of up-to-date materials from which they can pick
and choose. So in much the same way that a computer application is compiled from
the contributions of many developers, \gp allows individual classes and entire
courses to be `compiled' from the contributions of many skilled teachers and
developers. As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016} of which the CS+X approach is perhaps best-known \citep[\eg][]{Mir2017}; however, \gp represents a significant advance in this area since it not only builds in context at the level of individual modules, but at the level of a shared set of teaching resources as well!

\section{System Architecture}\label{system-architecture}

At least in our experience, the pace of technological and methodological change
is such that the gap between what instructors know and what students know is
substantial, and in many cases might even be growing. The differences within
individual student cohorts may be greater still, however, and while students
need to be able to begin their learning `journey' at some arbitrary point in the
progression from `learning to code' towards environmental, geospatial, or computational social scientist \citep{Lazer2009}. Instructors require the ability to assemble a course that meets students where they \emph{are} rather than where the instructor or institution might \emph{wish} them to be. At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if the means of progression is overly structured and rigid\ldots or just too slow to keep their attention.

The guiding insight behind \gp is that instruction in computation proceeds from
fundamental units of learning typically built around \textit{computing} concepts
(variables, lists/arrays, dictionaries/hashes, functions/subroutines, etc.) to
fundamental units of learning built around \textit{analytic} concepts (cluster
analysis, point patterns, spatial autocorrelation, etc.). These units must then
be assembled in a way that speaks to the student cohort and its background; we
mean this in two ways: first, that the examples used must speak to budding
geographers, political scientists, historians, etc.; and second, that it should
be possible to develop courses for different cohorts without having to start
over from scratch. In other words, how can we enable teachers to reuse many of
the building blocks employed in an advanced course for Masters students in an
introductory undergraduate course?

\subsection{Components}\label{components}

From these constraints it was clear that our system needed to support a
`bottom-up' approach to instructional design. So although the development of a
course or class could---and probably should---start with a clear set of learning
aims and outcomes, at a certain point the instructor will be searching for a
examples with which to teach a particular concept: What \emph{is} a list or
dictionary? What is \textit{k}-means clustering? We settled on the term `Atoms'
to refer to these basic instructional units and, much like entries in the atomic
table, we felt that they could be grouped together into sets of related
concepts: the fundamentals of programming, point pattern analysis, machine
learning, etc.

\begin{wraptable}{r}{0.6\textwidth}
\centering
\caption{Overview of System Components}
\label{table:1}
\begin{tabularx}{0.6\textwidth} {sb}
 \toprule
 \textbf{Concept} & \textbf{Usage} \\
 \midrule
 Atom  & A `teachable idea' combining explanation and code for a core, typically singular, concept (\eg Variables, Lists, Projections, Robust Rescaling\ldots). \\
 Sessions  & A `teachable unit' combining \textbf{multiple atoms} for delivery in a particular learning context (\eg Beginner Lab-based Course, Intermediate Flipped Module, Online Course\ldots) \\
 Modules  & A \textbf{sequence of sessions} designed to achieve one or more pedagogical objective and/or required for credit in a particular education context. \\
\bottomrule 
\end{tabularx}
\end{wraptable}

Each atom would employ \emph{domain-specific} illustrative examples and code so as to anchor learning in problems and applications relevant to the learner. An Atom on arrays written for geographers could start by showing how these simple data structures can be used to hold data about cities (\eg name, country, population); but a subsequent unit could build on this with an an illustration of how nested data structures allow us to add city location using nested coordinate pairs (\eg latitude and longitude within an array-of-arrays). This is why managed collaboration around sets of related concepts is so important: quite aside from the fact that no \textsc{foss} project succeeds without some level of executive control, it's critical that \textit{someone} is responsible for ensuring a degree of consistency in how the units fit together within a set of closely-related topics.

The purpose of the bottom-up approach is that these units can then be flexibly assembled into Sessions: from the \textit{same ingredients} (\ie Atoms) the instructor could create two quite different modules---one for delivery in person, one for delivery online---by organising the elements in different ways. Quite simply, we don't want to have to rewrite material for each format: lecture/practical, `flipped' classroom, or online course. Furthermore, sessions designed for experts (\eg those pursuing continuing online education) might be able to `move' students through many more atoms of instruction in a single session than a similarly laid out (in terms of time) session designed for first-time programmers in an undergraduate programme. So sessions need to be able to \emph{incorporate} atoms in a way that minimises the level of effort involved in finding the `best' way to, for example, explain the concept of recursion while maximising the ability of the instructor to relate this concept to the students' practical experience (\eg by providing a `context' that is anchored in a locally relevant `story' or data).

Naturally, Sessions can then be grouped into Learning Modules (modules) that offer a coherent instructional programme over a period of weeks or months. Modules represent the highest level of abstraction in the proposed system, but they are also obviously the starting point from which instructors can organise their (remixed) atomic and sessional material into something incorporating a set of learning outcomes and a package of assessment appropriate to their students. However, our design reflects the expectation that contributions to \gp at each \emph{level} of instruction might be made by \emph{different} people: a domain expert in Spatial Bayes might be the right person to develop an Atom on the concept and its application, but not the right person to develop a module tackling advanced spatial analytic concepts where this is just one approach amongst many. Similarly, given the global diversity of delivery formats, a 10-week term in Britain enables students to cover a very different `volume' of
content from a 15-week American semester. \gp recognises and seeks to respond to
that diversity at all levels.

\subsection{Tools}\label{tools}

So we are trying to design a system in which sessions and modules are
\emph{composed} out of atomic units of instruction that can be, optionally,
surrounded by the instructor's own `narrative' or supported by more locally
relevant (geo)data. We therefore want to produce a set of teaching materials
that are highly portable, easily re-used or edited, and that enable the
instructor to select only the elements from which they wish to compose their
materials. As intimated above, \gp operationalises this through the Jupyter
project and its ability to provide in-browser access to the Python
interpreter\footnote{Other languages are also possible using different kernels;
there is no reason that \gp couldn't be used to create instructional materials
in these languages as well.}, and it also takes advantage of the dominance of
the Git tool---and the GitHub service/web site---for version control as a means for tracking authorship across many versions and changes.

In theory, thanks to the combination of Jupyter and GitHub is not even necessary
for the novice user to have Python installed on their own computer: since all
interaction with Python is via the browser, the environment could be hosted on a
server halfway round the world. In practice, however, there are few such
services and most users simply download and install a free version of Python
(\eg Anaconda) that will run on their system. In our field many people are
already using this approach: notebooks can be found covering everything from
introductory concepts \citep{millington_reades_2017_code} to advanced spatial
analysis methods \citep{darribas_gds15}, and combined for both complete courses
or workshops \citep{rey_2016_narsc}.

Jupyter notebooks are written as a mix of executable code cells and
non-executable text formatted with the widely-used `markdown' syntax. Notebook
structure is provided through headers in markdown cells: a `\#' pre-pended to a
line of text is generally taken to be the title of the notebook; `\#\#' at the
start of a line provides a second level of structure (\ie level 2 headers);
`\#\#\#' indicates level 3 headers; etc. For our purposes, what's relevant is
that these headers naturally yield a semantic hierarchy that corresponds closely to the h1\ldots{}h6 model used by the HTML markup language that lies at the heart of the World Wide Web. This hierarchy allows us to `abstract out' the problem of inferring the \emph{meaning} of different sections of the notebook since the instructor does it for us through their use of headers.

\subsection{Approach}

In order to compile Atoms into Sessions and Modules, \gp necessarily requires a compositional syntax. We've noted the conceptual mapping between markdown and \textsc{html} formatting, above but how do we select some mix of code and markdown material in one notebook to be incorporated into another? And how do we do this in a way that is both simple to express and able to resolve ambiguity? Fortunately, such a model already exists and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets (\textsc{css}) uses well-understood `selectors' to specify one or more elements on a web page to which a set of presentational styles should be applied.

With \textsc{css} an `h1' in a style sheet indicates that all \textsc{html} Level 1 Headers should observe the styling rules declared immediately afterwards; while `h1.important' specifies that only a Level 1 Header of the class `important' should be selected. In fact, \textsc{css} also allows for nested selectors: for some element(s) of a `larger' selection to be selected in turn. This is normally used to do things like specify mouseover behaviours for a menu: that all anchors (\ie links) that are within a division of class \texttt{menu} should act in \textit{this} way when the mouse passes over them. What is particularly elegant about this approach is that it provides a means for selecting and styling multiple pieces of content in the document in one declaration (where this is desirable) \textit{and} a means for disambiguating content with the same name but in different locations within a document's hierarchical structure (where it is not).

Conceptually, \gp reframes this in terms of the meaning imparted by the instructor to the notebook: all cells coming after a Level 1 Header are considered to be part of that element's semantic field until another Level 1 Header is encountered or the end of the document is encountered, whichever comes first. So a Level 2 Header coming `after' (or `below', if you prefer) a Level 1 Header is still part of that higher-level element's semantic field, \textit{but} we can select it uniquely within the notebook using the standard \textsc{css} form of \texttt{h1.content h2.subcontent}. And with this we have essentially repurposed \textsc{css} as a means of selecting and importing content from one notebook into another! This is illustrated in schematic form in Figure \ref{fig:structure} where the \texttt{>} imply a hierarchical relationship.

\begin{wrapfigure}{r}{0.51\textwidth}
  \centering
  \caption{Illustration of Atom Structure}
  \includegraphics[width=0.5\textwidth]{Single_Document.pdf}
  \label{fig:structure}
\end{wrapfigure}

Unfortunately, the nature of Jupyter notebooks does not allow this to happen dynamically at run-time, but it does allow something similar to happen when an instructor is `compiling' new Sessional and Module content. In short, the instructor writes whatever content they wish but, using syntax similar to the examples below, wherever they want to incorporate contributed content from \gp (or elsewhere) they have only to `include' it by specifying both a source and a selection. And this approach also works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots

\subsection{Syntax}\label{syntax}

Adopting the approach outlined above: we typically envision an Atom as a short notebook focussing on a core concept or method (\eg lists, object-oriented design, or spatial autocorrelation); some or all of each Atom can then be selected and imported into a Session, which is itself a notebook; and the sessions can then be selected and managed through a Module, which can be a notebook or a set of notebooks. This process is initiated by the instructor creating a blank text cell in a Jupyter notebook and writing an `include' statement. The statement should be the \textit{only} content in the cell since \gp will be replacing the cell with an unknown number of whole text and code cells from the referenced notebook.

Crucially, \texttt{include} statements can be freely intermingled with the instructor's own content, allowing the instructor to `frame' the concepts in a way that suits their teaching style but which saves them having to reinvent the wheel for each class. A Session tackling standardisation could include elements of the relevant Atom from \gp while still allowing the instructor to interject comments, observations, questions, and additional tasks to ground the learning experience in the local context (individual, institutional, etc.). To illustrate this more clearly, an Atom on Python's approach to dealing with lists could be incorporated into a longer Session as follows:

\begin{Verbatim}[fontsize=\small]
@include {
    'nb'     = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
    'select' = 'h1.Lists'
}
\end{Verbatim}

\texttt{nb} is a path---local or remote---to a valid Jupyter notebook from which the Session or Module developer wants to import content. The \texttt{select} parameter specifies a selector for which the \gp tool will search within the source notebook. All content from that point onwards \emph{up to the next selector at he same level} will the then be copied into the compiled notebook. In example above, if there were a following \texttt{h1} covering, for example, `List Operators' then this would \emph{not} be included because, from a structural standpoint, it is of equal importance (at the same level in the hierarchy) to `Lists' but has \emph{not} been selected. Furthermore, any \texttt{h2} or \texttt{h3} subsections within the `Lists' section \textit{would} be included since they are presumed to be providing pedagogical and logical structure to the Understanding Lists section and so should be carried over.

Clearly, an instructor might want to import only part of of a section, or to
suppress a subsection falling in the middle of a larger resource. In
anticipation of this need, more complex `include' statements with no equivalent
in \textsc{css} are also possible:

\begin{Verbatim}[fontsize=\small]
@include {
    'resource' = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
    'select' = 'h1.Lists -h3.Lists Example; h1.List Operators -h2.Concatenation'
}
\end{Verbatim}

In this second example, two level 1 section are imported at the same time (the
selections are separated with semi-colons) and a level 3 subsection from
\emph{within} each of those sections is suppressed using the `-' syntax to
indicate that the section should be removed. An additional difference from true
\textsc{css} is that we allow spaces in the `selector' because we felt that
asking teachers to translate between a natural language header (``Understanding
Lists'') and what \textsc{css} would consider a safe header
(``Understanding\_Lists'') would detract from ease-of-use. A second departure
from the \textsc{css} standard is that selectors are not separated with commas:
we wanted to allow for this punctuation to be part of a section heading and felt
that semi-colons are rather more rare in that context.

\subsection{Putting it All Together}

Notebooks are actually stored in a format called JavaScript Object Notation
(\textsc{json}) that is not particularly easy for most humans to read, but as it
is nonetheless highly-structured we can interact with it programmatically. The
extensible nature of the \textsc{json} format also allows us to read and write
both data and metadata not only to each notebook, but also to each and every
sub-element of a notebook. Since metadata that is not understood by Jupyter is
simply ignored, we can add our own fields to provide useful information related
to instruction: the amount of time that the notebook's examples and content
might take to cover in class; who should be given credit for contributing; any
dependencies or requirements for installed libraries\ldots

In short, within the notebook we have two types of structure that can be used
not only to provide information that may assist course instructors in developing
content, but also to uniquely specify a section, subsection, or sub-subsection
of content (code or text) at a fine-grained level. Taken together, this provides
the foundation for a mechanism to select content in one notebook and import it
into another. This therefore creates an opportunity to remix/mash up content
while still being able to add an institutional or course-specific gloss wherever
necessary. Each notebook might start with the instructor's contact information
or by providing instructions for setting up the computing environment, but then
make use of material developed by others for actual instruction. This process
may seem quite abstract---and probably quite convoluted as well---but an
illustration (Figure \ref{fig:Illustration} \vpageref[above]{fig:Illustration})
may help to clarify why this process is so useful.

\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth, angle=0]{Multiple_Documents.pdf}
  \caption{Illustration of \gp Notebook `Compilation' Process}
  \label{fig:Illustration}
\end{figure}

Finally, in order to manage all of this output we make use of Git and the GitHub
web platform to monitor, approve, and roll back alterations to any submitted
revisions to \gp Atoms, Sessions, or Modules. This provides additional
capabilities such as allowing an instructor to `fix' their course to a
particular version of an Atom (a `release' or `tag' in Git) such that
development of the Atom by others can continue without the instructor having to
worry that the explanation or code upon which they rely will suddenly change!
Git also gives us the contributor information that we can propagate into
compiled notebooks so that credit is retained, and users and institutions can
better-understand the extent to which \gp is a \emph{joint} project relying on
the contributions of many teachers and developers.

\section{Use Cases}\label{uses}

Two use-cases: an online module of eight sessions, and an in-person module
lasting one semester might serve to further illustrate how this approach offers
a substantively new way to think about teaching programming material more
generally.

\subsection{An Online Module}\label{an-online-module}

There has been increasing interest at universities in Massive Open Online
Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education
through online delivery. For managers in Higher Education (\textsc{he}), the \textsc{mooc} promises both
enhanced revenue---depending on the model this implies either more people paying
to take the modules or more paying for a certificate of completion at the
end---and enhanced access as people who would be unable to attend the university
in person are nonetheless able to pursue a degree. Our own experience suggests
substantial student interest in `computational social science'
\citep{Lazer2009}, with international students being the most keen on such
modules as they are seen to provide marketable skills (programming) for
graduates from a discipline (geography) with generally high employment rates
(see \citeauthor{rgs2017} \citeyear{rgs2017}).

Whatever the motivation, while expanding access to, and enrolment by, students
is probably a `good thing', the process for developing a new online module or
`converting' an existing face-to-face module to an online format remains far
from easy. Typically, this step requires significant effort and investment by
both the instructor and their institution: all course materials need to be
online, recordings/videos of the lecturer presenting a key concept need to be
made, the sessions need to be `chunked' so that a 3-hour practical becomes
something more manageable, and assessments often need to be substantively
redesigned.

\gp significantly reduces the overhead of several of these stages: rather than
focussing on tasks such as how to explain or illustrate a particular concept,
the instructor can focus the \emph{majority} of their effort on constructing the
type of richer, fuller narrative that \textsc{mooc}s seem to require to engage
students working remotely. If the instructor isn't there to ask questions and
probe student learning on the spot, then the `frame' for each session (as well
as for the module as a whole) becomes crucial as it helps to anchor student
learning in concrete outcomes and encourages them to reflect upon, and
integrate, their progress over the self-guided learning process.

\subsection{In-Class Delivery}\label{in-class-delivery}

In a more traditional in-person format, \gp could be employed differently: since
the instructor is able to interactively provide the `frame' or `scaffold' (see
relevant discussion of `hypermedia' in \citealp{Azevedo2008}) upon which the
learning is built, there is less need for a narrative around each task or weekly
session. Here, the instructor might simply import a group of atoms to cover a
set of concepts and step through them interactively with the students. The
session could wrap up with a targeted project or mini-assessment that requires
the students to translate the concepts into a new problem domain or investigate
the process in more detail: \emph{e.g.} ``we've seen how we can use pandas and
bokeh to analyse demographic groups in London, here's a link to the open data
store page for Phoenix with similar data\ldots{}'' or ``Do you think that
socioeconomic class is a useful metric? Investigate the source of the data we've
just used and see if it squares with your expectations\ldots{}''

In effect, courses built on top of \gp become mash-ups, with the author able to
mix and match atoms as needed for the module objectives, time available, and
level of student. Both sessions and modules are built via composition: atoms are
composed into sessions, sessions into modules. The role of the instructor is
then to provide an integrative narrative that, in a sense, explains their choice
of components. In both the online and offline contexts we think that this has
the potential to free up instructors to concentrate on where they can most
effectively `add value', not in figuring out yet \emph{another} way to explain
how a list or dictionary works, but in explaining why they matter to a
geography, political science, or literature major.

\section{Engagement}\label{engagement}

Because it builds on \textsc{foss} approaches to software development, we expect \gp to benefit from network effects: the more people use it, the more useful it becomes, and the more people use it. The open source, peer-generated approach would, however, also be likely to present something of a challenge over time: as more people seek both to use and to contribute to \gp we would be likely to see the emergence of divergent norms, examples, and data across Atoms, Sessions, and Modules. In other words, the coherence of \gp as a teaching platform would become a victim of any real success in terms of widespread adoption.

\subsection{Coordination}

Here we believe that the \pysal project offers a useful template. Although \pysal is a free open source project, it is \emph{not} a free-for-all: domain specialists tend to gravitate towards those parts of the project to which they have the most to contribute and, over time, those who coordinate and enable the most substantive contributions to the codebase in terms of features and performance tend to become the `managers' of individual components of the overall tool (\eg lib, model, explore, viz). Overall project coherence is maintained via regular calls and online discussion boards, and a synchronised release schedule so that changes can be coordinated and knock-on effects identified.

We expect \gp to follow a similar path in that a relatively small subset of users will contribute the majority of the content; however, unlike a traditional software project there is an important role for \emph{teachers}, not just highly-skilled \emph{developers}. We think that this represents a really exciting opportunity for innovative new approaches to rise to the surface. There may be only a few who can write the Python code to conduct a Geographically-Weighted Regression analysis, but it will be interesting to see how many creative, insightful ways there are to explain it!

The educational focus implies that it may well be the most committed teachers who end up coordinating and reviewing contributions to \gp from other instructors and developers. We would also argue for the relevance of grouping sets of Atoms together in basic libraries such as `foundations' (the basics of variables and data structures); `description' (describing data); etc. Within these sets, contributions would be managed by a coordinator to ensure that examples, style, and other features are consistent for use in a set of Sessions introducing Python programming, or Local Indicators of Spatial Autocorrelation.

The key to \gp is that it seeks to balance the benefits of code-sharing with those of local expertise: the instructor is free to write their own exegesis, if you will, of the code and its relevance to a particular Session or Module, but the burden of developing a cogent, domain-specific demonstration can be shared with others and compelling examples more widely adopted without the effort of reinventing the wheel. Where irreconcilable differences arise between pedagogical approaches or models, then we might expect to see small groups of collaborators `fork' the codebase and offer their own models; this is, of course, a valid approach with open source code and one from which all instructors can ultimately benefit!

\subsection{Credit}

\gp also acknowledges the reality of the need for peer and professional recognition by incorporating attribution mechanisms through which faculty can be credited for their work. Contributors can be recognised as the author of content elements through a Digital Object Identifier (\textsc{doi}), and The \gp team would also include section editors responsible for the curation of atoms and for maintenance of official \gp \textsc{doi}s. Instructors who use \gp to compose sessions or courses can also contribute those modules back to the project and would also be recognised as the author of record in a similar fashion. Authorship of \gp educational components, be they atoms, sessions or modules, provides the contributors with peer-evaluated, impactful materials to add to their promotion applications.

Although the open source nature of the project means that we cannot strictly
enforce attribution, \gp seeks to make this the easier to do `by default'
through the use of Git to track and insert contributory metadata into the
compiled notebook. We can use this metadata to append a list of contributors to
the end of each notebook, along with any other relevant acknowledgments or
copyright notices. To facilitate re-use while protecting contributions from
unacknowledged exploitation textual content in \gp is covered by a Creative
Commons license; however, to deal with the fact that \gp relies extensively on
open source contributions which are incompatible with some \textsc{cc} licenses
\citep[see discussion in][]{osswatch2013}, code blocks are licensed under the
\textsc{mit} license. The manner in which contributions from authors at
different institutions can be combined also `pollutes' the materials in ways
that inhibit institutional assertions of ownership over \gp content.

\section{Limitations}\label{limitations}

Although we are profoundly excited both by the possibility of shifting to
compositional approaches to content-development, and by the impact that shared
materials could have on our teaching, it is highly unlikely that our first
attempt to divide up the entire field into discrete units of instruction will be
entirely successful. For this, it will be useful to reference documents such as
the \textit{Body of Knowledge} \citep{bok2018} and \textit{Subject Benchmark
Statement} \citep{QAA2014} for the `why' and `what' of instruction, leaving \gp
to deal with the `how'. Moreover, the open source, contributory nature of the
project positions us to build \gp on top of the shared understanding of many
specialists with a range of ideas about how to break apart---and put back
together---the constituent elements of our domain knowledge in ways that speak
to different types of students.

For the time being we have also deliberately hobbled \gp in one important way:
an \texttt{include} command must be in a cell that does not contain any other
text or code. This was done primarily for simplicity: it's a lot easier to look
for whole (text) cells that match a target pattern than to have to try to parse
long blocks of text or code on the off-chance that something might match; it
also avoids any ambiguity as to whether the \texttt{include} is a \gp command or
`real' code as might happen, for instance, if \gp were used with a different
programming language. Not coincidentally, it is also a good deal easier to know
that you are replacing an entire block (the one with an include) with one or
more entire blocks that are a mix of text and code, than trying to work out
whether the block in which an include was found needs to be `closed out' before
the import starts.

\section{Conclusion \& Future Directions}\label{future}

From practical experience, conference presentations, and code we tend to already
know who is a good \emph{programmer} or \emph{theorist}, \gp provides a
mechanism for discovering who is a good \emph{teacher}. Sometimes these
abilities may reside in the same person, but more often we expect that they will
not: the strongest developers tend to be people who have been practicing
software development for many years and, consequently, may have difficulty
communicating their ideas to beginner- or intermediate-level students (or
teachers!). For this reason we see the broad-based community-of-practice aspect
of \gp as integral at all stages of the project: system enhancement, content
development, expanding coverage, and instructional design.

Consequently, \gp has a lot in common---both philosophically and
practically---with the Software Carpentry movement \citep{SCF2016}, and although
we seek to tackle a slightly narrower set of issues with a more re-usable set of
resources, we can take both inspiration and warning from their experience. The
benefit, we think, is that while it is possible to design \gp sessions and
modules that follow the popular `bootcamp' approach to instruction (though see
critique in \citealp{Feldon2017}), we want to enable the \textit{same} content
to be employed in a flexible carpentry format as well as a `normal' classroom or
\textsc{mooc} as required\ldots Or even to enable the instructor to mash all of
those formats together such that they use a `bootcamp' format for the
introduction to Unix and the command line, an online-formatted \gp resource for
foundational concepts in computer science, and a traditional course format for
the (geo)data analysis instruction. All pulled from the same source!

Ultimately, although \gp was developed with teaching needs in mind there is, of
course, no reason why it couldn't be put to other uses: in combination with with
\texttt{nteract} (\url{https://github.com/nteract/nteract}) it would allow
developers or researchers to build fully-fledged applications as scripts
assembled from a collection of notebooks; or as an addition to Netflix's
notebook ecology to allow for enhanced resource-sharing and standardisation
during the development phase before features and interfaces are `fixed' as
libraries \citep{Ufford2018}. Nonetheless, our focus for the time being remains
the cohort of teachers at all levels tasked with introducing programming
material to their students and wondering where to begin. We hope that \gp makes
a valuable contribution to this application domain and look forward to working
with others to roll out a rich, reusable teaching framework.

\bibliographystyle{apacite}
\bibliography{Geopyter.bib}

\end{document}